### 2016-05-22 tex/toy.tex
recast the story in a cleaner way

\documentclass[preprint,3p]{elsarticle}
%\documentclass[preprint]{aastex}
\usepackage{aas_macros}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\DeclareMathOperator\erf{erf}
\newcommand{\mli}[1]{\mathit{#1}}
%\usepackage{epstopdf}

\begin{document}

%\begin{frontmatter}
%
%\title{Toy Model}
%\author{A.~G. Kim\corref{cor1}}
%\ead{agkim@lbl.gov}
%\address{Physics Division, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley CA, USA 94720}
%
%\begin{abstract}
%\end{abstract}
%\begin{keyword}
%\end{keyword}
%\end{frontmatter}

\section{Likelihood and Posterior}
\label{likelihood:sec}

The ambition is to construct a model that predicts the observations of
transients.
The conceptual structure of the model is shown in Figure~\ref{pgm:fig}. 
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{../results//hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   \label{pgm:fig}}
\end{figure}

There are top-level  sets of parameters $M$ that correspond to photometric calibration, cosmology,
host-galaxy properties, and transient properties.
There are two sets of intermediary parameters $M_I$ that correspond to luminosity $L$ and type $\tau$.

The model has the following types of observables $O$ that
may be observed for each transient:
\begin{itemize}
\item $c$: Counts of the photometric measurements (light curves).
\item $s$: Counts of the spectroscopic measurements.
\item $z$: Redshift.
\item $T$: Transient type.
\end{itemize}
The actual measurements of the observables are labeled  with a subscript ``o''.

We are interested in what the model predicts for the pdf of those observables
for the population of transients that enter the analysis.  
The probabilities for this ``analysis'' population are notated with the
subscript $S$, for example the likelihood is expressed as $\mathcal{L}_S(O| M, M_I)$.

The likelihood can be expressed in terms of probabilities of the observables
for all possible transients and observables accommodated by the model.
The probabilities for this ``underlying'' population are notated with no subscript,
so that the pdf of observables for this population is $p(O|M, M_I$).
The analysis population differs from the underlying population in that the transients of the former 
must, for the typed SN~Ia analysis, be discovered, pass photometric selection criteria, be selected for spectroscopic
observation, have a successful type from the spectrum, and be typed as SN~Ia.

The likelihood is proportional to the pdf of the underlying sample that
enter the analysis sample, as expressed by
\begin{align}
% \mathcal{L}_S(O_o | M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
\mathcal{L}_S(O | M, M_I) & \propto p(O, \text{in analysis}|M, M_I)\\
&  \propto \prod_{\alpha \in transients} p(O^\alpha, \text{in analysis}|M, M_I),
\end{align}
where the second line holds because the observations and stochastic decisions
for sample selection are expected to be independent from transient to transient.

For a single supernova labeled by $\alpha$,   the probability of the analysis population is related
to the underlying population by
\begin{align*}
p_S(O^\alpha|M, M_I) &\equiv p(O^\alpha, \text{in analysis}|M, M_I)\\
& = P(disc=True, anal=True, sobs=True, type=True, T^\alpha=SNIa|O^\alpha, M, M_I)p(O^\alpha|M, M_I)\\
& = P(disc=True, anal=True, sobs=True, type=True|T^\alpha=SNIa, O^\alpha, M, M_I)\\
& \times P(T^\alpha=SNIa| O^\alpha, M, M_I)p(O^\alpha|M, M_I)\\
&= P(disc=True|T^\alpha=SNIa,c^\alpha, M, M_I)
P(anal=True|T^\alpha=SNIa,c^\alpha, M, M_I) \\
&  \times P(sobs=True|T^\alpha=SNIa,c^\alpha, M, M_I)
 P(type=True|T^\alpha=SNIa, s^\alpha, M, M_I)\\
&  \times
P(T^\alpha=SNIa| M, M_I)p(c^\alpha|M, M_I)p(s^\alpha|M, M_I)p(z^\alpha|M, M_I).
\end{align*}
where we anticipate that the stochastic processes that enter each sample-selection step
are independent of each other, and that the measurement uncertainties are independent.
We find that $p_S$ is the pdf of the underlying
distribution $p(O^\alpha|M, M_I)$ multiplied by the efficiency of a realized target to
enter the analysis sample.

The likelihood can be expressed as
\begin{align}
\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients}p_S(O_o^\alpha|M, M_I)}{\int  \prod_{\alpha \in transients} p_S(O^\alpha |M, M_I)dO}.
\label{likelihood:eqn}
\end{align}
A frequentist analysis can proceed with this likelihood.
Alternatively, the Bayesian posterior is specified by parameter priors as 
\begin{align*}
p(M, M_i | O_o) & \propto \mathcal{L}_S(O_o| M, M_I) p(M_I|M) p(M)\\
 & \propto \mathcal{L}_S(O_o| M, L, \tau) p(L| \tau, M) p(\tau|M) p(M).
\end{align*}
Once all the $p$'s and $P$'s are specified we are ready to analyze!

%We now consider the terms in the likelihood contributed by each transient and omit the $\alpha$ superscript. 
%The dependencies of the parameters are charted through the PGM to give for transients with a  spectroscopic type measurement
%\begin{align*}
%p(O| M, M_I) & = P(T |\tau) p(z | M) p(c | L, M) p (s|L, M),
%\end{align*}
%and for transients without a type measurement (not relevant for the SN~Ia-only analysis but included for completeness)
%\begin{align*}
%p(O| M, M_I) & = p(z | M) p(c | L, M),
%\end{align*}
%where the photometric and spectroscopic observations have independent measurement uncertainties.
%

%Transient discovery and sample selection are purely based on observed photometry $c$.
%The decision for spectroscopic follow-up is also based on observed photometry,  though perhaps external redshift estimates
%of the possible host galaxy may come into play.  The success of the typing measurement should primarily depend
%on the signal $s$, but may also depend on the type and host properties.

\section{Probabilities: Minding Ones P's and p's}
Section~\ref{likelihood:sec} introduces the probabilities and probability distribution functions that describe the likelihood
and posterior for the model shown in Figure~\ref{pgm:fig}.
In this Section we present one possible model; the assembly of other models should
resemble this example.
We now drop the $\alpha$ label with the implicit understanding that the distributions
now described are per transient.

\subsection{$p(M)$: Priors on the Top-Level Parameters}
\begin{itemize}
\item
Flat $w$CDM cosmological parameters are drawn from top-hat distributions
\begin{align}
\Omega_M & \sim  {\mathcal{U}}(0,1)\\
w_0 & \sim \mathcal{U}(-1.5, -0.1).
\end{align}

\item
Photometric zeropoints for each band are given by
\begin{equation}
Z \sim \mathcal{N}\left(0,\left(\frac{\ln{10}}{2.5}\right)^2 C_{cal}\right),
\end{equation}
where $C_{cal}$ is the covariance of the calibration zeropoints calculated by Chris Lidman.
(The PGM anticipates a more complicated calibration approach.)

\item
There are a number of parameters that describe the supernova populations.

The luminosity of SNe~Ia as a function of time and wavelength is described by the SALT2 model and
the linear relation of its parameters with absolute magnitude.
The following priors are taken from the Rubin et al.\ UNITY paper:
\begin{align}
M_{SNIa} & \sim \mathcal{U}(-20, -18) \\
\arctan{\alpha_{SNIa}} & \sim \mathcal{U}(-0.2, 0.3) \\
\arctan{\beta_{SNIa}} & \sim \mathcal{U}(-1.4, 1.4),
\end{align}
and for ``hyperparameters'' that describe the distributions of the per-supernova parameters
\begin{align}
\log_{10}{R^{x_0}_{SNIa}} & \sim \mathcal{U}({-2}, {-0.5})\\
\log_{10}{R^{x_1}_{SNIa}} & \sim \mathcal{U}({-0.5}, {0.5})\\
\log_{10}{R^{c}_{SNIa}} & \sim \mathcal{U}({-1.5}, {-0.5})\\
x_{1,SNIa}^\star& \sim \text{Cauchy}(0,1)\\
c^\star_{SNIa} & \sim \text{Cauchy}(0,0.3).
\end{align}

The following are irrelevant for the cosmology analysis of the pure SN~Ia sample
but are presented for completeness.
We take the rates relative to the SN~Ia rate
\begin{equation}
r_{SNIa} = 1.
\end{equation}
If we were interested in the absolute rates this term could be replaced with your favorite parameterized
SN rate model.
The relative non-Ia rate would be a stochastic parameter, with a prior
\begin{equation}
r_{SNII} \sim \mathcal{U}(0.25, 4).
\end{equation}
In a better model this relative rate would be redshift-dependent.
The description of non-Ia models is deferred.

\item 
In the photometric-sample analysis a prior must be specified for the redshift.  The redshift is drawn from a flat distribution
\begin{equation}
z\sim \mathcal{U}(0,\infty).
\end{equation}
\end{itemize}

\subsection{$p(M_I|M)$: Probabilities of the Intermediary Parameters.}
\begin{itemize}
\item Type probabilities are relevant in the photometric analysis.
The probability an object is SN~Ia is
\begin{align}
P_{SNIa} &= \frac{r_{SNIa}}{r_{SNIa}+r_{SNII}} \nonumber \\
P_{SNII}&=1-P_{SNIa}.
\label{prob:eqn}
\end{align}

The type of the object $\tau$ is drawn from the Bernoulli distribution 
\begin{equation}
\tau | p_{SNIa} \sim \mathcal{B}(1-p_{SNIa}),
\end{equation}
with the convention that $T=0$ corresponds to SN~Ia.

There are parameters that specify the sub-type.
\begin{align}
t_0 & \sim \mathcal{U}\\
x_{0, SNIa} & \sim \mathcal{N}(0,R^{x_0}_{SNX})\\
x_{1,SNIa} & \sim \mathcal{N}(x_{1,SNIa}^\star,R^{x_1}_{SNIa})\\
c_{SNIa} & \sim \text{SkewNormal}(c^\star_{SNIa},R^{c}_{SNIa} )
\end{align}

\end{itemize}

\subsection{$p(O| M, M_I)$: Probability of All Possible Observables}
\begin{itemize}
\item The spectroscopic typing is assumed to be perfect
\begin{equation}
T | \tau \sim \delta (T-\tau).
\label{typepdf:eqn}
\end{equation}

\item The spectroscopic redshift of a spectroscopically typed transient assumed to be perfect
\begin{equation}
z | z_{host} ~ \delta (z-z_{host}).
\end{equation}

When the transient is not spectroscopically typed, there is ambiguity in the correct assignment of the host galaxy.
Then the measured redshift is a sum of delta functions
\begin{equation}
p(z_o|z) = \sum_i p_i \delta(z-z_i),
\end{equation}
where galaxy $i$ at
redshift $z_i$ has probability $p_i$ of being
the host.

\item  The luminosity distance comes from the physics-theory function
\begin{equation}
d_L = d_L(z; \Omega_M, w0),
\end{equation}
the flux for an observation at date $t$ is
\begin{equation}
f = 10^{-0.25 \left(M_{SNIa} - \alpha x_1 + \beta c \right)}\frac{1}{4\pi d_L^2} F_{SALT2}((t-t_0)/(1+z); x_0,  x_1, c, \lambda/(1+z)),
\end{equation}
and expected counts is
\begin{equation}
c_\star = 10^{Z/2.5}f.
\end{equation}
The distributions of counts are 
\begin{align*}
c | c_\star &\sim \mathcal{N}(c_\star, C_c)\\
s | c_\star &\sim \mathcal{N}(c_\star, \sigma_s),
\end{align*}
where $C_c$ and $C_s$ are the measurement covariances for the photometric and spectroscopic observations respectively.

\end{itemize}

\subsection{Sample Selection Probability}
We have listed the terms that contribute to the probability that a target enters
the analysis sample.

\begin{itemize}
\item $P(T=SNIa| M, M_I) = \delta_{\tau, SNIa}$: The transient is
measured to be SN~Ia if it is SN~Ia.
\item $ P(type=True|T=SNIa, s, M, M_I)$: 
The probability of successful spectroscopic typing needs to be determined.
A simple possibility is that  it only depends on the signal-to-noise of the spectroscopic observation
$\epsilon_T(ston)$.  Then given a  spectroscopic noise of $\sigma_s$
\begin{equation}
P(type=True|  T=SNIa, s, M, M_I) = \epsilon_T(s/\sigma_s).
\end{equation}

\item $P(disc=True|T=SNIa,c, M, M_I)$:
The probability for discovery of a transient in a single image depends on the
signal-to-noise $\epsilon_D(ston)$.  The probability of discovery given all
the images is
\begin{equation}
P(disc=True | T=SNIa,c,  M, M_I)
=  1-\prod_{\beta \in images} \left(1-\epsilon_D(c_\beta/N_\beta)\right).
\end{equation}

\item $P(anal=True|T=SNIa,c, M, M_I)$:
As an illustrative example we write out the probability of a analysis-sample selection requiring two photometric points above
a signal-to-noise threshold $STON_T$: other sample selection cases are similar but with different combinatorics
\begin{align*}
P(anal=True  | T=SNIa, c,  M, M_I)= &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\\
&  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)),
 \end{align*}
 where $\mathcal{H}$ is the Heaviside step function.

\item $P(sobs=True|T=SNIa,c, M, M_I)$:
The probability a transient has a spectroscopic observation.
Mat Smith and Chris D'Andrea should be able to provide this term.
Until we get this, lets make the simplifying ansatz that
\begin{align*}
P(sobs=True | T=SNIa, c, M, M_I)= & const.
 \end{align*}

\end{itemize}

\section{Implementation Notes}
Anticipating the use of available statistical packages, the likelihoods have been constructed
from standard distribution functions.  The truncated distributions from sample selection
are accounted by the normalization 
terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
in the PGM, they are treated as nodes who contribute 
$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
factor.

The following integral appears in the normalization
\begin{equation}
\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
\end{equation}
We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
\begin{equation}
n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
\end{equation}
the integral becomes
 \begin{equation}
\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
\end{equation}
The two probability distributions are normal so that this reduces to
 \begin{equation}
\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
\end{equation}
where
\begin{equation}
c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
\end{equation}

\end{document}