### 2018-12-20 papers/methods/misc/author_list_arxiv.tex

% Author list file generated with: mkauthlist.py 1.2.4 
% mkauthlist.py -j arxiv --sort-nonbuilder -a order.csv author_list.csv author_list_arxiv.tex 
S. R. Hinton, T. M. Davis, A. G. Kim, D. Brout, C. B. D'Andrea, R. Kessler, J. Lasker, C. Lidman, E. Macaulay, A. M\"oller, M. Sako, D. Scolnic, M. Smith, R. C. Wolf, M. Childress, E. Morganson, S. Allam, J. Annis, S. Avila, E. Bertin, D. Brooks, D. L. Burke, A. Carnero Rosell, M. Carrasco Kind, J. Carretero, C. E. Cunha, L. N. da Costa, C. Davis, J. De Vicente, D. L. DePoy, P. Doel, T. F. Eifler, B. Flaugher, P. Fosalba, J. Frieman, J. Garc\'ia-Bellido, E. Gaztanaga, D. W. Gerdes, R. A. Gruendl, J. Gschwend, G. Gutierrez, W. G. Hartley, D. L. Hollowood, K. Honscheid, E. Krause, K. Kuehn, N. Kuropatkin, O. Lahav, M. Lima, M. A. G. Maia, M. March, J. L. Marshall, F. Menanteau, R. Miquel, R. L. C. Ogando, A. A. Plazas, E. Sanchez, V. Scarpine, R. Schindler, M. Schubnell, S. Serrano, I. Sevilla-Noarbe, M. Soares-Santos, F. Sobreira, E. Suchyta, G. Tarle, D. Thomas, V. Vikram, Y. Zhang (DES Collaboration)